### 2017-10-22 paper/sections/4_methodology.tex
paper: fix microbenchmark section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: space reduction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply    & apply to metadata store in obj store \\
  Nonvolatile Apply & apply to metadata store in memory \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{Mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

@@

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability ~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{append client journal+RPCs} since
both mechanisms do the same thing or \texttt{stream+save} since ``global"
durability is stronger and has more overhead than ``local" durability.

@@

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. To implement DeltaFS on Cudele, the user
would use the configuration:

\begin{listing}[h]
\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "append_client_journal"
    "durability": "local_persist"
}
\end{minted}
\end{listing}

BatchFS merges updates into the global namespace after the job is complete.
This can be achieved with by composing two mechanisms:

\begin{listing}[h]
\begin{minted}[xleftmargin=21pt,fontsize=\footnotesize,
               tabsize=4]{js}
{     
  "allocated_inodes": "100000"
  "interfere_policy": "block"
  "consistency": "append_client_journal+volatile_apply"
  "durability": "local_persist"
}
\end{minted}
\end{listing}

>>>>

\begin{table}
\begin{tabular}{ r | l }
  Mechanism         & Description \\\hline
  RPCs              & round trip remote procedure calls \\
  Stream            & stream journal into object store \\
  Append Client Journal & events appended to in-memory journal \\
  Volatile Apply & apply to metadata store in memory \\
  Nonvolatile Apply    & apply to metadata store in obj store \\
  Local Persist     & journal saved to client's disk \\
  Global Persist    & journal saved in object store \\
\end{tabular}
\caption{Mechanisms are composed together to form consistency and
durability semantics.\label{table:mechanisms}} 
\end{table}

@@

% which system they represent and which are impossible
Existing, state-of-the-art systems in HPC can be represented by the cells in
Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
have global consistency and durability~\footnote{This is the normal case.
IndexFS also has bulk merge which would transition the system into ``weak
consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
and BatchFS uses ``weak" consistency and ``local" durability. These systems
have other features that could push them into different semantics but we assign
labels here based on the points emphasized in the papers.  To compose the
mechanisms users inject which mechanisms to run and which to use in parallel
using a domain specific language.  Although we can achieve all permutations of
the different guarantees in Table~\ref{table:spectrum}, not all of them make
sense. For example, it makes little sense to do \texttt{append client
journal+RPCs} since both mechanisms do the same thing or \texttt{stream+save}
since ``global" durability is stronger and has more overhead than ``local"
durability. The cost of each mechanism and the semantics described above are
quantified in Sections~\S\ref{sec:microbenchmarks} and~\S\ref{sec:use-case-1}.

@@

% examples
Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation.  BatchFS merges updates into the global
namespace after the job is complete.  This can be achieved with by composing
two mechanisms:

### 2017-10-22 paper/sections/6_results.tex
paper: space reduction
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Cudele lets users construct consistency/durability guarantees using
well-established research techniques from other systems; so instead of
evaluating the scalability and performance of the techniques
themselves\footnote{ This is the reason we did not compare to other HPC or data
center systems; we are not claiming that our file system metadata protocols are
superior.}, we show that (1) the mechanisms we propose are useful for
constructing semantics used by real systems, (2) the techniques can work
side-by-side in the same namespace, and (3) the combination of these techniques
can help the system scale farther when subtrees are coupled to the correct type
of application.  We scope the evaluation to one metadata server and scale the
clients, which execute an extremely metadata intensive job in parallel. This
type of analysis shows the capacity and performance of a metadata server with
superior metadata protocols, which should be used to inform load balancing
across a metadata cluster. Load balancing with partitioning and replication can
be explored with something like Mantle~\cite{sevilla:sc15-mantle}.

\begin{table}
\begin{tabular}{ r | l | l}
  Cluster  & Hardware               & Software \\\hline
  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
\end{tabular}
\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
monitor daemon, MDS is metadata server daemon, and OSDs are object storage
daemons.\label{table:clusters}} 
\end{table}

We evaluate our prototype on the two clusters in Table~\ref{table:clusters}.
Microbenchmarks were run on the in-house clusters while client scalability
tests were run on CloudLab. We only scale up to 29 clients because that is the
maximum capacity of a single metadata server. For the in-house cluster, we need
more object storage servers and we have to co-locate clients on the same node
because of resource contention \footnote{The bandwidth to RADOS is only
sufficient with 8 nodes on this older hardware, which uses SSDs and nodes that
are almost 10 years old.  The CloudLab cluster performs just as well with and
without a highly provisioned object storage server cluster.}. All daemons run
as a single process which is the default setting for Ceph. We develop on Ceph's
Jewel release, version 10.2.1, which was released in May 2016.  We run
experiments on the same cluster and normalize results to control for hardware
differences and to make our results more generally applicable.

This paper adheres to The Popper
Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}, so
experiments presented here are available in the repository for this
article\footnote{Links have been removed for double-blind submission}.
Experiments can be examined in more detail, or even re-run, by visiting the
\texttt{[source]} link next to each figure. That link points to a Jupyter
notebook that shows the analysis and source code for that graph, which points
to an experiment and its artifacts.

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{ [\href{https://...}{source}] Microbenchmark: performance of each
mechanism (left) and building consistency/durability guarantees (right) for
100K files on a single client. Results are normalized to the runtime of writing
files creates to the client's in-memory journal.
\label{fig:composable-mechanisms}}
\end{figure}

@@

{\it Journal Size:} We measure the amount of storage per journal update to be
about 2.5\text{KB}. The storage footprint scales linearly with the number of
metadata creates and suggests that updates for a million files would be
\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
payloads of this size in most HPC/data cente networks are reasonable.\\

\noindent\textbf{Takeaway}: measuring the mechanisms individually shows that
their overheads and costs can differ {\it by orders of magnitude}. Cudele gives
users the ability to quantify the performance impact of different
consistency/durability guarantees.\\

@@

@@

% Weak Consistency benefit not due to metadata format
{\it Metadata Formats:} Because the metadata formats are the same for
all schemes we argue that the performance gain for decoupled namespaces comes
from relaxing the consistency guarantees and not from the metadata formats.\\

\noindent\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.

@@

\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
\caption{ [\href{https://...}{source}] Use Case 2: using the ``allow" and
``block" API, users can isolate directories from interfering clients. Variance
with blocking turned on is the same as ``isolated" from
Figure~\ref{fig:overhead-b}.
\label{fig:slowdown-allow-block}}

\end{figure}

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
Next we show how Cudele can be programmed to block interfering clients. We
use the same problematic workload from Figure~\ref{fig:overhead-b}, where
clients write to their own private directories and another client interferes with a stream of creates at
30 seconds.  We also have another client write to a decoupled namespace and
merge its updates at 90 seconds.  We equip each client directory with the
configuration:

\begin{listing}[tb]
\begin{minted}[xleftmargin=21pt,
               tabsize=4]{js}
{     
    "allocated_inodes": "100000"
    "interfere_policy": "block"
    "consistency": "RPCs"
    "durability": "stream"
}
\end{minted}
\end{listing}

@@

% results
To show the benefits of this isolations, our results in
Figure~\ref{fig:slowdown-allow-block} are plotted alongside the variance bars from
Figure~\ref{fig:overhead-b}. Because ``allow"/``interfere and
``block"/``isolated" have the same variability we draw the following three
conclusions: (1) clients that use the API to block interfering clients  get
the same performance as isolated clients, (2) there is a negligible effect on
performance for the extra work the metadata server does to return
\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.\\

\noindent\textbf{Takeaway}: the API lets users isolate directories when
applications need better and more reliable performance. This is a way of
controlling consistency.

@@

\begin{figure}[tb] \centering
\includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png} \caption{
[\href{https://...}{source}] Use Case 3: slowdown of a single client periodically syncing
updates to the gobal namespace. The crossover point represents the trade-off of
frequent updates (with small journal files) and infrequent updates (with larger journal
files).\ref{fig:slowdown-sync}.  \label{fig:slowdown-sync}}
\end{figure}

\begin{figure}[tb] \centering
\includegraphics[width=1.0\linewidth]{graphs/mergescale.png} \caption{
[\href{https://...}{source}] Use Case 3: performance of merging client journals
at the metadata server. Performance is better than RPCs because (1) there are
less messages on the network and (2) the consistency/durability code paths in
the metadata server are bypassed. Merges are serialized and read from the local
disk, so keeping client journals in memory or on SSD will improve the
performance of ``decoupled" even more.  \ref{fig:mergescale}.
\label{fig:mergescale-sync}}
\end{figure}

@@

To show the contention at the metadata server of this approach, we scale the
number of journal merge events, each hitting different directories in the
namespace, in Figure~\ref{fig:mergescale}. The size of each journal is 100K
files, the maximum recommended directory size for CephFS. We also plot the cost
of RPCs with and without journaling; recall from Figure~\ref{fig:overhead-c}
that the scalability that max number of clients that the default metadata
server could support with RPCs is 18.

@@

Performance is better than RPCs because we place no restrictions on the
validity of metadata inserted into the journal and avoid touching poorly
scaling data structures. In other words, we can scale linearly if we weaken our
consistency by never checking for the existence of files before creating files.
only appends \texttt{open()} requests to the journal. When the updates are
merged by the metadata server into the in-memory metadata store they never scan
growing data structures.  Had we implemented the ``Create" mechanism to include
\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
the poor scaling that we see with the ``RPCs" mechanism.\\

\noindent\textbf{Takeaway}: the weakest form of consistency for creating files
({\it i.e.} not checking data structures for the validity of an update or
existence of a file) shows linear scalability and stable performance.

\subsection{Use Case 4: Reading Large Directories}

% CITEME
The final use case is reading large directories. At job completion, scientists
might use \texttt{ls} again to see the results. This causes great strain on the
file system as paths need to traversed and the entire directory, with all its
entries, must be transferred back to the client. Here we show how decoupling a
large namespace and materializing the view in memory on the client is faster
than doing RPCs for walks of the file system namespace.

FigureX shows how fast a client can materialize decoupled namespaces in memory.
This takes the on-disk file system metadata format and parses them into events
that can be manipulated and appended to. Again, we scale the number of up

\subsection{Discussion}
\begin{itemize}
  \item Composable guarantees: applications control their performance (knobs you didn't have before)
  \item Hardware features: adding hardware in strategic places
  \item Performance Prediction: quantifies the cost of each mechanism
\end{itemize}

>>>>

Cudele lets users construct consistency/durability guarantees using
well-established research techniques from other systems; so instead of
evaluating the scalability and performance of the techniques
themselves\footnote{This is the reason we did not compare to other HPC or data
center systems; we are not claiming that our file system metadata protocols are
superior.}, we show that (1) the mechanisms we propose are useful for
constructing semantics used by real systems, (2) the techniques can work
side-by-side in the same namespace, and (3) the combination of these techniques
can help the system scale further when subtrees are coupled to the correct type
of application.  We scope the evaluation to one metadata server and scale the
clients, which execute an extremely metadata intensive job in parallel. This
type of analysis shows the capacity and performance of a metadata server with
superior metadata protocols, which should be used to inform load balancing
across a metadata cluster. Load balancing with partitioning and replication can
be explored with something like Mantle~\cite{sevilla:sc15-mantle}.

%\begin{table}
%\begin{tabular}{ r | l | l}
%  Cluster  & Hardware               & Software \\\hline
%  In-House & 15 nodes, 4 2GHz CPUs  & Ubuntu 12.04, 1 MON\\
%           & 8GB RAM, SSD           & 1 MDS, 8 OSDs, 8 Clients\\
%  CloudLab & 34 nodes, 16 2.4GHz CPUs & Ubuntu 14.04, 1 MON\\
%           & 128GB RAM, SSD         & 1 MDS, 3 OSDs, 29 Clients
%\end{tabular}
%\caption{In-House used for microbenchmarks, CloudLab for Use Cases. MON is
%monitor daemon, MDS is metadata server daemon, and OSDs are object storage
%daemons.\label{table:clusters}} 
%\end{table}

We run experiments on the same cluster and normalize results to control for
hardware differences and to make our results more generally applicable.  We use
a CloudLab cluster of 34 nodes, each with 16 2.4 GHz CPUs, 128GB RAM, and SSDs.
Each node uses Ubuntu 14.04 and has 1 monitor server, 3 object storage daemons,
1 metadata server, and 29 clients.  All daemons run as a single process which
is the default setting for Ceph. We develop on Ceph's Jewel release, version
10.2.1, which was released in May 2016.

This paper adheres to The Popper Convention~\cite{jimenez_popper_2016}, so
experiments presented here are available in the repository for this article.
Experiments can be examined in more detail, or even re-run, by visiting the
\texttt{[source]} link next to each figure. That link points to a Jupyter
notebook that shows the analysis and source code for that graph, which points
to an experiment and its artifacts.

\subsection{Microbenchmarks}
\label{sec:microbenchmarks}
\begin{figure}[tb]
\centering
\includegraphics[width=1.0\linewidth]{graphs/composable-mechanisms.png}
\caption{[\href{https://www.google.com}{source}] Microbenchmark: performance of each
mechanism (left) and building consistency/durability guarantees (right) for
100K files on a single client. Results are normalized to the runtime of writing
files creates to the client's in-memory journal.
\label{fig:composable-mechanisms}}
\end{figure}

@@

{\it Journal Size:} We measure the amount of storage per journal update to be
about 2.5\text{KB}. The storage footprint scales linearly with the number of
metadata creates and suggests that updates for a million files would be
\(2.5\text{KB}*1\texttt{ million files} = 2.38\text{GB}\). Transfer times for
payloads of this size in most HPC/data cente networks are reasonable.

\textbf{Takeaway}: measuring the mechanisms individually shows that
their overheads and costs can differ {\it by orders of magnitude}. Cudele gives
users the ability to quantify the performance impact of different
consistency/durability guarantees.\\

@@

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/mergescale.png}
      \caption{Use Case 1} \label{fig:mergescale}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/block-allow.png}
      \caption{Use Case 2}
      \label{fig:block-allow}
  \end{subfigure}
  \begin{subfigure}[b]{.3\linewidth}
      \centering
      \includegraphics[width=1.0\linewidth]{graphs/slowdown-sync.png}
      \caption{Use Case 3}
      \label{fig:slowdown-sync}
  \end{subfigure}
  \caption{The performance and features of Cudele. (a) shows the performance of
merging client journals at the metadata server. Performance is better than RPCs
because there are less messages and consistency/durability code paths are
bypassed. (b) shows how the allow/block API can isolate directories from
interfering clients.  (c) shows the slowdown of a single client periodically
syncing updates to the gobal namespace. The inflection point is the trade-off
of frequent updates vs. larger journal files.\label{fig:use-cases}}
\end{figure*}

@@

% Weak Consistency benefit not due to metadata format
{\it Metadata Formats:} Because the metadata formats are the same for
all schemes we argue that the performance gain for decoupled namespaces comes
from relaxing the consistency guarantees and not from the metadata formats.

\textbf{Takeaway}: we confirm the performance benefits of other
well-established research systems but, more importantly, we show that the
Cudele mechanisms we propose are useful for building and evaluating different
consistency/durability semantics.

@@

% setup: 2cs sep dirs, 2cs DN, 1c malicious write
Next we show how Cudele can be programmed to block interfering clients. We use
the same problematic workload from Figure~\ref{fig:overhead-b}, where clients
write to their own private directories and another client interferes with a
stream of creates at 30 seconds.  We also have another client write to a
decoupled namespace and merge its updates at 90 seconds.  We equip each client
directory with the configuration: allocated\_inodes=100K,
interfere\_policy=block, consistency=RPCs, and durability=stream.

@@

% results
To show the benefits of this isolations, our results in
Figure~\ref{fig:allow-block} are plotted alongside the variance bars from
Figure~\ref{fig:overhead-b}. Because ``allow"/``interfere and
``block"/``isolated" have the same variability we draw the following three
conclusions: (1) clients that use the API to block interfering clients  get
the same performance as isolated clients, (2) there is a negligible effect on
performance for the extra work the metadata server does to return
\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
negligible effect on performance.

\textbf{Takeaway}: the API lets users isolate directories when
applications need better and more reliable performance. This is a way of
controlling consistency.

@@

@@

To show the contention at the metadata server of this approach, we scale the
number of journal merge events, each hitting different directories in the
namespace, in Figure~\ref{fig:mergescale}. The size of each journal is 100K
files, the maximum recommended directory size for CephFS. We also plot the cost
of RPCs with and without journaling; recall from Figure~\ref{fig:overhead}
that the scalability that max number of clients that the default metadata
server could support with RPCs is 18.

@@

Performance is better than RPCs because we place no restrictions on the
validity of metadata inserted into the journal and avoid touching poorly
scaling data structures. In other words, we can scale linearly if we weaken our
consistency by never checking for the existence of files before creating files.
only appends \texttt{open()} requests to the journal. When the updates are
merged by the metadata server into the in-memory metadata store they never scan
growing data structures.  Had we implemented the ``Create" mechanism to include
\texttt{lookup()} commands before \texttt{open()} requests, we would have seen
the poor scaling that we see with the ``RPCs" mechanism.

\textbf{Takeaway}: the weakest form of consistency for creating files
({\it i.e.} not checking data structures for the validity of an update or
existence of a file) shows linear scalability and stable performance.

%\subsection{Use Case 4: Reading Large Directories} 
%
%% CITEME
%The final use case is reading large directories. At job completion, scientists
%might use \texttt{ls} again to see the results. This causes great strain on the
%file system as paths need to traversed and the entire directory, with all its
%entries, must be transferred back to the client. Here we show how decoupling a
%large namespace and materializing the view in memory on the client is faster
%than doing RPCs for walks of the file system namespace.
%
%FigureX shows how fast a client can materialize decoupled namespaces in memory.
%This takes the on-disk file system metadata format and parses them into events
%that can be manipulated and appended to. Again, we scale the number of up

%\subsection{Discussion}
%
%Cudele and the experiments we present here prompt many avenues for future work.
%The first is embeddable policies, where child subtrees have specialized
%features but still maintain the guarantees  of their parent subtrees. We
%already show how applciations can compose guarantees and control their
%performance, knobs that we did not have before, but embeddable policies are a
%way to compose the policies themsleves.  For example, a RAMDisk subtree is
%POSIX-compliant but relaxes durability constraints, so it can reside under a
%POSIX subtree that is strongly consistent. These embeddable policies should be
%controlled and enforced by Cudele.
%
%Second is the opportunity for adding hardware in strategic places to speed up
%the mechanisms that the application needs. In Figure~\ref{fig:mergescale} the
%performance of merging updates at the metadata server is limited by the speed
%of the disk. So applications that rely on merging many updates should consider
%equipping the metadata server with SSDs or more memory. This is a form of
%co-design that takes a ``dirty-slate" approach.
%
%Finally, performance prediction benefits from our cost quantification of each
%mechanism. Applications can use Cudele to microbenchmark their components and
%software, similar to what we did in Secion~\S\ref{sec:microbenchmarks}. Using
%those results, they can predict how much slower their system will be if they
%adopt stronger consistency or durability. This can also be a verification tool
%where performance that varies wildly from the predicted performance can be a
%red flag that something is wrong or that the bottleneck is not in the
%consistency or durability plane.
